---
title: "final project"
name: "Freya Huang"
output: html_document
date: "2023-12-03"
---
```{r -1, warning=FALSE, echo=FALSE}
library(readxl)

ny4 <- read_excel("~/Desktop/Research/dataset/ny_april_10009_HAN (1) (1) 2.xlsx")
head(ny4)
```

```{r 0, warning=FALSE,echo=FALSE}
nyp <- read_excel("~/Desktop/Research/dataset/ny_property_HAN (1) (1) 2.xlsx")
head(nyp)
```


#combine two datasets
```{r 1, warning=FALSE,echo=FALSE}
library(tidyverse)
nyp4 <- left_join(ny4, nyp, by = c('PropertyID' = 'Property ID'))
head(nyp4)
```

#create dataset nyd
#prepared the data

```{r 2, warning=FALSE,echo=FALSE}
nyd_ <- nyp4[,c(-1,-3,-6:-12,-15:-20,-22,-32,-43:-44,-52,-53,-56,-57)]

#change "NA" to 0
nyd_[,20] <- as.numeric(nyd_[[20]])
sum(is.na(nyd_[,20]))
nyd_[is.na(nyd_[, 20]), 20] <- 0
sum(is.na(nyd_[,20]))

nyd_[,21] <- as.numeric(nyd_[[21]])
nyd_[is.na(nyd_[, 21]), 21] <- 0
sum(is.na(nyd_[,21]))

nyd_[,25] <- as.numeric(nyd_[[25]])
nyd_[is.na(nyd_[, 25]), 25] <- 0

#remove useless listing types
nyd_1 = nyd_[nyd_[,5] != "40827",]
nyd_1 = nyd_1[nyd_1[,5] != "41330",]
nyd_1 = nyd_1[nyd_1[,5] != "41935",]

#convert char to num variables 
nyd_1[,12] = as.numeric(nyd_1[[12]])
nyd_1[,16] = as.numeric(nyd_1[[16]])
nyd_1[,17] = as.numeric(nyd_1[[17]])
nyd_1[,22] = as.numeric(nyd_1[[22]])
nyd_1[,25] = as.numeric(nyd_1[[25]])
nyd_2 = na.omit(nyd_1)


#convert char to factor
nyd_3 <- nyd_2 %>% 
  mutate(
    `Property Type` = as.factor(`Property Type`),
    Neighborhood = as.factor(Neighborhood),
    `Listing Type` = as.factor(`Listing Type`),
    Superhost = as.factor(Superhost),
    `Cancellation Policy` = as.factor(`Cancellation Policy`),
    `Business Ready` = as.factor(`Business Ready`),
    `Instantbook Enabled` = as.factor(`Instantbook Enabled`)
  )

# dependent variable setting for nyd
nyd_4 <- nyd_3 %>%
  mutate(B = ifelse(Status == "B", 1, 0))
nyd_sta <- nyd_4[,-2]

# Assuming your dataset is named 'your_data'
# Replace 'your_data' with the actual name of your dataset

# Load the dplyr package for additional functionality
library(dplyr)

# Basic summary statistics
summary_stats <- nyd_sta %>%
  summarise(
    Number_of_Rows = n(),
    Number_of_Columns = ncol(.),
    Missing_Values = colSums(is.na(.))
  )

head(nyd_sta)

# summary statistics
#number of rows: 9180(observations)
#number of colomns: 34
#missing value: 0

```

# Summary the daily blocking status for each property
```{r 3, warning=FALSE,echo=FALSE}
library(dplyr)
library(ggplot2)

nyd_sta <- nyd_sta %>%
  mutate(DailyRevenue = `Average Daily Rate`)

# Calculate the fraction of the number of days of a month already supplied (always 30 days)
nyd_sta <- nyd_sta %>%
  mutate(FractionDaysSupplied = `Count Available Days LTM` / 30)
```

```{r 4, warning=FALSE,echo=FALSE}
library(tidyverse)

nyd_sta_processed_ <- nyd_sta %>%
  group_by(PropertyID) %>%
  summarise(
    TotalDays = n(),
    BlockedDays = sum(B == 1),
    FractionSupplied = BlockedDays / TotalDays
  )
#The summary for each unique PropertyID in nyd_sta. The summary includes the total number of days (TotalDays), the number of days where the condition B == 1 is true (BlockedDays), and the fraction of days supplied under this condition (FractionSupplied).

# Join the processed data with the original dataset
nyd_sta_processed_ <- nyd_sta_processed_ %>%
  left_join(nyd_sta, by = "PropertyID")

nyd_sta_processed <- nyd_sta_processed_ %>%
  group_by(PropertyID) %>%
  slice(1)

nyd_sta_processed <- nyd_sta_processed %>%
  select(-B)

# Calculate the fraction of days supplied
nyd_sta_processed <- nyd_sta_processed %>%
  mutate(FractionDaysSupplied = `Number of Bookings LTM` / 30)
nyd_sta_processed$BlockingStatus <- nyd_sta_processed$BlockedDays / nyd_sta_processed$TotalDays

head(nyd_sta_processed)
```



# Create a ggplot for Supply Curve Analysis:
```{r,echo=FALSE}
library(ggplot2)
nyd_sta_plot <- ggplot(nyd_sta_processed, aes(x = FractionSupplied, y = `Average Daily Rate`, color = BlockingStatus)) +
  geom_point() +
  labs(x = "Blocking Status", y = "Average Daily Rate", title = "Supply Curve Analysis") +
  scale_color_gradient(low = "blue", high = "red", name = "Blocking Status") +
  theme_minimal()

summary(nyd_sta_plot)
nyd_sta_plot
```

# Perform linear regression
```{r,echo=FALSE}
library(dplyr)
lm_model <- lm(`Average Daily Rate` ~ FractionSupplied, data = nyd_sta_processed)

# Get regression line equation and R-squared value
regression_equation <- as.character(lm_model$coefficients[1]) %>%
  paste(" + ", as.character(lm_model$coefficients[2]), " * x")

r_squared <- summary(lm_model)$r.squared

# Print the results
cat("Regression Equation: y =", regression_equation, "\n")
cat("R-squared Value:", r_squared, "\n")
```



## according to different Neighborhood and property types
#All Neighborhood Analysis:
```{r,echo=FALSE}
plot_all <- ggplot(nyd_sta_processed, aes(x = FractionSupplied, y = Price)) +
  geom_point(aes(color = as.factor(`Property Type`), shape = as.factor(Neighborhood))) +
  labs(x = "Blocking Status", y = "Daily Revenue") +
  scale_color_discrete(name = "Property Type") +
  scale_shape_discrete(name = "Neighborhood")
plot_all
```
#East Village Neighborhood Analysis:
```{r,echo=FALSE}
plot_EastV <- ggplot(nyd_sta_processed, aes(x = FractionSupplied, y = Price)) +
  geom_point(data = subset(nyd_sta_processed, Neighborhood == "East Village"),
             aes(color = as.factor(`Property Type`))) +
  labs(x = "Blocking Status", 
       y = "Daily Revenue") +
  scale_color_discrete(name = "Property Type")
plot_EastV
```

#Stuyvesant Town Neighborhood Analysis:
```{r,echo=FALSE}
plot_Stown <- ggplot(nyd_sta_processed, aes(x = FractionSupplied, y = Price)) +
  geom_point(data = subset(nyd_sta_processed, Neighborhood == "Stuyvesant Town"),
             aes(color = as.factor(`Property Type`))) +
  labs(x = "Blocking Status", 
       y = "Daily Revenue") +
  scale_color_discrete(name = "Property Type")
plot_Stown 
```

# write the dataset
```{r,echo=FALSE}
library(writexl)

# Export to Excel
write_xlsx(nyd_sta, "nyd-sta.xlsx")
```

